Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-01 paper/paper.tex
pass on technique

\Mar{----------------------------- estou aqui}

\tname{} uses a composition of heurisitc functions to sort the rules
it produces. Each function uses a different criterion to estimate how
similar a rule, created in the previous step, is to rules from a
public database~\cite{emerging-threats-open} including a total of
\numrulessuri\ rules. We describe in the following each of these
heuristic functions. The codomain of each function ranges in the 0-1
interval. The score of a rule is the average across the values
produced by each function.

@@

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity between a
synthesized and a human-created rule is the size of the synthetic
rule, \ie{}, the number of options it contains. Informally, rules with
popular sizes score high whereas inpopular rules score
low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes for our rule data set. The x-axis shows the number of
options included in a rule (up to 10) while the y-axis shows the
number of rules with the corresponding number of options. Note that
there are 11 rules with no options. These rules only use the header to
signal a traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

%\tname{} uses this information to rank rules.

More precisely, we calculate the contribution of size of a given rule
$r$ with the following formula:

\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]

\noindent
, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
dictionary mapping sizes to number of associated rules, and
$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
for any dictionary entry. Consequently, the contribution of size is
proportional to its popularity. For example, the contribution of this
heuristic for a rule with three options would be 0.684 (=3,662/5,353)
whereas the contribution of a rule with nine options would be 0.028
(=154/5,353). Note that the measurement is normalized in the 0-1
range.

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} \Luc{92\%} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]

\noindent
Note that the only difference of this formula compared to the formula
to compute contribution of size is the distribution used, which is
given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
a dictionary of number of content options to number of rules that
contain that number of content options.

\subsubsection{Popularity of content strings.} 
This heuristic weighs higher rules that use popular strings in content
options. For a given rule \tname{} takes the arithmetic mean of the
popularity scores of each string used in contents in that rule. More
precisely, we calculate the contribution of content rarity with the
following formula:

>>>>

\tname{} uses a composition of heurisitic functions to sort the rules
produced in the previous step. Each of these functions uses a
different criterion to estimate how similar a created rule is to those
from public rulesets~\cite{emerging-threats-open}. The
intuition is that rules that look more similar to existing trusted
rules rank higher and could catch the attention of a system
administrator. The score of a rule is the weighted average across the
values produced by each heuristic function. The codomain of each
function ranges on the 0-1 interval. Hence, the score of a rule also
ranges on the 0-1 interval. We describe in the following each of these
heuristic functions.

@@

\subsubsection{Size of a rule}
\label{sec:heuristic-size-of-rule}
One element that \tname{} uses to measure similarity between a
synthesized and a human-created rule is the size of the synthetic
rule, \ie{}, the number of options it contains. Informally, rules with
popular sizes score high whereas unpopular rules score
low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes for our rule data set. The x-axis shows the number of
options included in a rule (up to 10) while the y-axis shows the
number of rules with the corresponding number of options. Note that
there are 11 rules with no options. These rules only use the header to
signal a traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes. We calculate the contribution of size of a given rule
$r$ with the following formula $\mathit{rules\_per\_size}[\mathit{len(r)}]/\mathit{max(rules\_per\_size)}$, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
$\mathit{rules\_per\_size}$ encodes the distribution of sizes as a
dictionary mapping sizes to number of associated rules, and
$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
for any dictionary entry as to normalize the measurement in the 0-1
range. Consequently, the contribution of size is proportional to its
popularity. The contribution of this heuristic would be 0.684
(=3,662/5,353) to a rule with three options and 0.028 (=154/5,353) to
a rule with nine options.

%% \vspace{-2ex}
%% \[\]
%% \vspace{-2ex}
%\noindent

%% \Luc{Se for preciso temos numeros para fundamentar isso
%%   melhor, sao cerca de 70K contents nas 28K regras do
%%   data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
%%   (\%) das 28K regras tem contents.} \Luc{92\%}

@@

\subsubsection{Number of content options.} We observed
that \percContentOptions\ of the rules from the ruleset we are
analyzing contain content options, but the distribution of those
contents per rule is not uniform. Figure~\ref{} shows the
corresponding distribution. Given the importance of this kind of
option to the definition of rules, we defined a heuristic function
focused on the typical number of content options in a rule. We
calculate the contribution of this function as described on previous
Section~\ref{sec:heuristic-size-of-rule}.

%% to the score of a rule
%% with the following formula:

%% \vspace{-2ex}
%% \[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
%% \vspace{-2ex}

%% \noindent
%% The only difference of this formula compared to the one defined on
%% Section~\ref{sec:heuristic-size-of-rule} is the distribution used. The
%% distribution is given by the term
%% $\mathit{rules\_per\_number\_of\_contents}$ denoting a dictionary of
%% number of content options to number of rules that contain that number
%% of content options.

\Mar{------------------------------------------- estou aqui}
\subsubsection{Popularity of content strings.} 
This heuristic weighs higher rules that use popular strings in content
options. For a given rule \tname{} takes the arithmetic mean of the
popularity scores of each string used in contents in that rule. More
precisely, we calculate the contribution of content rarity with the
following formula:
